\documentclass[a4paper, 10pt, oneside]{memoir}

% Essentials
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[hidelinks]{hyperref}
\usepackage[backend=bibtex,style=ieee]{biblatex}
\usepackage[tmargin=3cm, bmargin=2cm, lmargin=2cm, rmargin=2cm]{geometry}
\usepackage[backend=bibtex,style=ieee]{biblatex}

% Bibliography
\bibliography{main}

% Optional but useful
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{float}
\usepackage{subfiles}
\usepackage{enumitem}
\usepackage{tabularx,booktabs,multirow}
\usepackage{longtable}
\usepackage{url}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{mdframed}
\usepackage{csquotes}
\usepackage[table]{xcolor}
\usepackage{titling}

% Memoir
\setcounter{secnumdepth}{3}


% Indentation
\setlength\parindent{0pt}

% Title
\title{Genetic Algorithms: Summaries}
\date{\today}
\author{\textbf{Jonas~Devlieghere} \\ \texttt{jonas.devlieghere@student.kuleuven.be}}

% Theorems
\theoremstyle{definition}
\newtheorem{ex}{Example}[subsection]
\newtheorem{experiment}{Experiment}[subsection]

% Question item
\newcommand{\question}[2]{%
    \item {\textbf{#1}} \vspace{0.5em} \\  \textit{#2}
}
\newcommand{\unanswered}{\vspace{10em}}

\begin{document}

% Title page
\maketitle
\newpage

% Table of Contents
\tableofcontents
\newpage

% Main Matter

\part{Summaries}

\chapter*{Foreword}
This chapter summarizes the papers that are part of the course \emph{Genetic Algorithms and Evolutionary
Computing}\footnote{\url{https://onderwijsaanbod.kuleuven.be/syllabi/e/H02D1AE.htm}} at the University of Leuven. The
reader is assumed to have a basic understanding of genetic algorithms. Many papers, especially the older ones, tend to
reintroduce the concept. These sections have been omitted. Sections containing results have either been omitted or
summarized, depending on the completeness of the concluding sections.
\\\\
This work is strictly intended for educational purpose. All text in this document is either summarized, paraphrased or
copied verbatim from the original paper. I do not claim this document provides any original work. As a prospective
engineer, I have little knowledge of the law regarding copyright. I'm assuming this work falls under the ``fair use''
or equivalent policy.


\chapter{Evolving Inventions}
\subfile{subfiles/evolving_inventions.tex}

\chapter{Genetic Algorithms, Constraints, and the Knapsack Problem}
\subfile{subfiles/knapsack.tex}

\chapter{Evolving 3D Morphology and Behavior by Competition}
\subfile{subfiles/evolving_morphologie.tex}

\chapter{Evolution of Adaptive Behavior in Robots}
\subfile{subfiles/adaptive_robots.tex}

\chapter{A Knowledge-Intensive GA for Supervised Learning}
\subfile{subfiles/knowledge_intensive_ga.tex}

\chapter{Using GA For Supervised Concept Learning}
\subfile{subfiles/supervised_concept_learning.tex}

\chapter{Optimization of Control Parameters for Genetic Algorithms}
\subfile{subfiles/optimization_control_paramters.tex}

\chapter{Maintaining Healthy Population Diversity}
\subfile{subfiles/healthy_population_diversity.tex}

\part{Study Material}

\chapter{TSP}
\subfile{subfiles/tsp.tex}

\chapter{Questions}
\subfile{subfiles/questions.tex}


% Appendix
\newpage
\appendix
\printbibliography

\end{document}
